The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2012-08-20 script/en/chapters/cryptomethods.tex
Some small corrections in the CT Script

% HACK to fix warning "destination with the same identifier .. has already been used, ...":
\makeatletter \renewcommand{\thepage}{~\csname @arabic\endcsname \c@page~} \makeatother
\hypertarget{Kapitel_1}{}
\chapter{Encryption Procedures}
\label{Label_Kapitel_1}
(Bernhard Esslinger, Joerg-Cornelius Schneider, May 1999; Updates Dec. 2001, Feb. 2003,
June 2005, July 2007, January 2010, June 2012)

@@

    With {\bf CrypTool 2}\index{CrypTool 2} you can execute the following modern
    symmetric encryption algorithms (using in the Startcenter Templates
    \textbackslash{} Cryptography \textbackslash{} Modern \textbackslash{}
    Symmetric):\\
    AES, DES, PRESENT, RC2, RC4, SDES, TEA, TripleDES, Twofish.

@@

More details about the AES algorithms and the AES candidates of the last round
can be found within the online help of CrypTool\index{CrypTool}%
\footnote{%
      CrypTool 1 online help\index{CrypTool 1}: The index head-word {\bf AES}
      leads to the 3 help pages: {\bf AES candidates}, 
      {\bf The AES winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}.
  }.

% HACK to fix warning "destination with the same identifier .. has already been used, ...":
\makeatletter \renewcommand{\thepage}{\csname @arabic\endcsname \c@page} \makeatother
% --------------------------------------------------------------------------
\subsection{New results about cryptanalysis of AES}
\label{NeueAES-Analyse}
\index{Cryptanalysis}
\index{AES}

@@

There are also cipher challenges for asymmetric algorithms (please
see chapter \ref{nt:NoteFactorization})\footnote{%
  A wide spectrum of simple and complex crypto riddles are included
  in the ``Mystery Twister C3''~challenge.
  \index{Challenge}
  \index{Cipher challenge}
  \index{Crypto challenge}
  }
.

@@

% --------------------------------------------------------------------------
\newpage
% \vskip +40 pt
\begin{center}
\fbox{\parbox{15cm}{
    \emph{IETF\footnotemark:} \\
    There is an old saying inside the US National Security Agency (NSA):\\
    "Attacks always get better; they never get worse."
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  \url{http://tools.ietf.org/html/rfc4270}\index{IETF}\index{NSA}
  }
\section{Ciphers and cryptanalysis for educational purposes} \index{Cryptanalysis}

@@

% --------------------------------------------------------------------------
\section{Further details}

Beside the information you can find in the following chapters, in many other
books and on a good number of websites, the online help of 
CrypTool\index{CrypTool} also offers very many details about the 
symmetric and asymmetric encryption methods.

@@

The Sage module \texttt{crypto/block\_cipher/miniaes.py} supports Mini-AES to allow
students to explore the working of a modern block cipher.

>>>>

% HACK to fix warning "destination with the same identifier .. has already been used, ...":
\makeatletter \renewcommand{\thepage}{~\csname @arabic\endcsname \c@page~} \makeatother
\hypertarget{Kapitel_1}{}
\chapter{Security Definitions and Encryption Procedures}
\label{Label_Kapitel_1}
(Bernhard Esslinger, Joerg-Cornelius Schneider, May 1999; Updates Dec. 2001, Feb. 2003,
June 2005, July 2007, January 2010, June 2012)

@@

    With {\bf CrypTool 2}\index{CrypTool 2} you can execute the following modern
    symmetric encryption algorithms (using in the Startcenter {\bf Templates
    \textbackslash{} Cryptography \textbackslash{} Modern \textbackslash{}
    Symmetric}):\\
    AES, DES, PRESENT, RC2, RC4, SDES, TEA, TripleDES, Twofish.

@@

More details about the AES algorithms and the AES candidates of the last round
can be found i.e. within the online help of CrypTool 1\index{CrypTool 1}%
\footnote{%
      CrypTool 1 online help\index{CrypTool 1}: The index head-word {\bf AES}
      leads to the 3 help pages: {\bf AES candidates}, 
      {\bf The AES winner Rijndael} and 
      {\bf The Rijndael encryption algorithm}.
  }.

% HACK to fix warning "destination with the same identifier .. has already been used, ...":
\makeatletter \renewcommand{\thepage}{\csname @arabic\endcsname \c@page} \makeatother
% --------------------------------------------------------------------------
\subsection{Results about cryptanalysis of AES}
\label{NeueAES-Analyse}
\index{Cryptanalysis}
\index{AES}

@@

There are also cipher challenges\footnote{%
  A wide spectrum of simple and complex crypto riddles are included
  in the ``MysteryTwister C3''~challenge: \url{http://www.mysterytwisterc3.org}.
  \index{Challenge}
  \index{Cipher challenge}
  \index{Crypto challenge}
  }  for asymmetric algorithms (please see chapter \ref{nt:NoteFactorization}).

@@

% --------------------------------------------------------------------------
\newpage
% \vskip +40 pt
\begin{center}
\fbox{\parbox{15cm}{
    \emph{IETF\footnotemark:} \\
    There is an old saying inside the US National Security Agency (NSA):\\
    "Attacks always get better; they never get worse."
}}
\end{center}
\addtocounter{footnote}{0}\footnotetext{%
  \url{http://tools.ietf.org/html/rfc4270}\index{IETF}\index{NSA}
  }
\section{Cryptanalysis and ciphers for educational purposes} \index{Cryptanalysis}

@@

% --------------------------------------------------------------------------
\section{Further information}

Beside the information you can find in the following chapters, in many other
books and on a good number of websites, the online help of all
CrypTool variants\index{CrypTool} also offer very many details about the 
symmetric and asymmetric encryption methods.

@@

The Sage module \texttt{crypto/block\_cipher/miniaes.py} supports Mini-AES to allow
students to explore the inner working of a modern block cipher.

### 2012-08-20 script/en/chapters/introduction.tex
Some small corrections in the CT Script

The later editorial work in TeX added cross linkages between different sections and footnotes describing where you can find the according functions within
the CrypTool programs. The appendices \ref{s:appendix-menu-overview-CT1} etc. contain a complete overview
of\index{CrypTool 1.x}\index{CrypTool 2}\index{JCrypTool}
the \hyperlink{menu-overview-CT1}{CrypTool 1 menu tree},
the \hyperlink{template-overview-CT2}{CrypTool 2 templates} and
the \hyperlink{function-overview-JCT}{JavaCrypTool (JCT) functions}.

@@

The script was delivered the first time with CrypTool\index{CrypTool} in
version 1.2.01. Since then it has been expanded and revised in almost
every new version of CrypTool 1.

@@

This document is delivered together with the open-source program
CrypTool 1\index{CrypTool 1.x}.

@@

CrypTool 1 (CT1)\index{CrypTool} is an educational program with a comprehensive
online help enabling you to use and analyse cryptographic procedures within a
unified graphical user interface.

CrypTool 1\index{CrypTool 1.x} and the successor versions
CrypTool 2 (CT2)\index{CrypTool 2.0} and JCrypTool (JCT)\index{JCrypTool 1.0}
are used world-wide for training in companies and teaching at schools and
universities.
\par \vskip + 15pt

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to CT1\index{CrypTool}. They applied their
very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Staff of Prof.\ Claudia Eckert, Prof.\ Johannes Buchmann,
         Prof.\ Torben Weis and Prof.\ Arno Wacker.
\end{list}
%\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\
\\
Bernhard Esslinger
\\
\\
Frankfurt (Germany), July 2012

>>>>

The later editorial work in TeX added cross linkages between different sections and footnotes describing where you can find the according functions within
the CrypTool programs. The appendices \ref{s:appendix-menu-overview-CT1} etc. contain a complete overview
of\index{CrypTool 1}\index{CrypTool 2}\index{JCrypTool}
the \hyperlink{menu-overview-CT1}{CrypTool 1 menu tree},
the \hyperlink{template-overview-CT2}{CrypTool 2 templates} and
the \hyperlink{function-overview-JCT}{JCrypTool (JCT) functions}.

@@

The script was delivered the first time with the CrypTool-1 package in
version 1.2.01. Since then it has been expanded and revised in almost
every new version of CrypTool 1.

@@

This document is delivered together with the open-source program
CrypTool 1\index{CrypTool 1}.

@@

CrypTool 1 (CT1) is an educational program with a comprehensive
online help enabling you to use and analyse cryptographic procedures within a
unified graphical user interface.

CrypTool 1\index{CrypTool 1} and the successor versions
CrypTool 2 (CT2)\index{CrypTool 2.0} and JCrypTool (JCT)\index{JCrypTool 1.0}
are used world-wide for training in companies and teaching at schools and
universities.
\par \vskip + 15pt

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to CrypTool\index{CrypTool}. They applied their
very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Hr.\ Dominik Schadow
   \item Staff of Prof.\ Claudia Eckert, Prof.\ Johannes Buchmann,
         Prof.\ Torben Weis and Prof.\ Arno Wacker.
\end{list}
%\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\
\\
Bernhard Esslinger
\\
\\
Frankfurt (Germany), July 2012

### 2012-08-20 script/en/chapters/paper_and_pencil.tex
Some small corrections in the CT Script

\item {\bf Double column transposition} \cite{pp:Savard1999}
   \index{Double column transposition}: 
   Double columnar transposition
   was frequently used during WW2 and during the Cold War. Two simple columnar
   transpositions with different keys are executed successively\footnote{%
   Using CrypTool: Choose a key for the 1st permutation, input line by line, 
   permute and output column by column. Then choose a (different) key for the
   2nd permutation, input line by line, permute and output column by column.}.

>>>>

\item {\bf Double column transposition} \cite{pp:Savard1999}
   \index{Double column transposition}: 
   Double columnar transposition
   was frequently used during WW2 and during the Cold War. Two simple columnar
   transpositions with different keys are executed successively\footnote{%
   Using CrypTool: Choose a key for the 1st permutation, input line by line, 
   permute and output column by column. Then choose a (different) key for the
   2nd permutation, input line by line, permute and output column by column.}.\\
   If the keys are different and long enough (at least each 20 characters),
   then this is even for today's computer a real challenge\footnote{%
     MTC3 offers an according level-X challenge called "Double Columnar
     Transposition":\\
     \url{http://www.mysterytwisterc3.org/en/crypto-challenges/level-x-cryptography-challenges?id=12}                                                     }.

### 2012-08-20 script/en/script-en.tex
Some small corrections in the CT Script

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Crypto Procedures}
\listof{cryptoprocedure}{List of Crypto Procedures}
\endgroup

@@

\vspace{30pt}
\begin{itemize}
  \item All samples have been tested with Sage v 4.7 (release date 2011-05-23).
  \item The source code of the Sage samples in this script is delivered as
        Sage program files within the CrypTool setup program.
        After installing CrypTool 1.x \index{CrypTool 1.x} you find
        them within the subdirectory \verb#sage# within the CrypTool directory:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

>>>>

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Crypto Procedures with Pseudo Code}
\listof{cryptoprocedure}{List of Crypto Procedures with Pseudo Code}
\endgroup

@@

\vspace{30pt}
\begin{itemize}
  \item All samples have been tested with Sage v 4.7 (release date 2011-05-23).
  \item The source code of the Sage samples in this script is delivered as
        Sage program files within the CrypTool-1 setup program.
        After installing CrypTool 1 \index{CrypTool 1} you find
        them within the subdirectory \verb#sage# within the CrypTool-1
        directory:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}
